---
title: "Emotional Intelligence as a Predictor of Academic Success"
author: "Joash"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(dplyr)
library(readr)
library(ggplot2)
source("../config/config.R", chdir = FALSE)
```

## Background

This report investigates whether self-reported emotional intelligence (EI)
indicators, augmented with NLP-derived features from open-ended student
reflections, can predict academic performance (GPA).

## Data

```{r load-data}
features <- read_csv("../data/features.csv", show_col_types = FALSE)
glimpse(features)
```

## NLP Features

The pipeline extracts three NLP-derived features from each student's free-text
reflection:

- `sentiment_compound` — overall sentiment via the `syuzhet` lexicon.
- `emotional_vocab_count` — number of words drawn from a curated emotion lexicon.
- `response_length` — token count, used as a control.

## Regression Model

```{r model}
fit <- lm(
  reformulate(EI_PREDICTORS, response = TARGET_VAR),
  data = features
)
summary(fit)
```

## Visualizations

```{r ei-vs-gpa}
ggplot(features, aes(x = empathy_score, y = gpa_score)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Empathy vs GPA", x = "Empathy", y = "GPA") +
  theme_minimal()
```

## Discussion

The headline correlation reported in the project README (r = 0.68) reflects
results obtained on the full study dataset, which is not redistributed in this
repository. With the small synthetic sample shipped here the regression is
illustrative only — readers should reproduce the pipeline on their own data
to verify effect sizes.

## Reproducibility

All analysis is driven from `analysis/01_data_cleaning.R` →
`analysis/02_nlp_processing.R` → `analysis/03_regression_analysis.R` →
`analysis/04_visualization.R`. Configuration lives in `config/config.R`.
